curiobaby_drop model analysis

### 2021-01-23 analysis/model_analysis_v2.Rmd

mdat <- mdat %>% rename(drop = drop_object,
                        target = target_object,
                        relation = condition) %>%
  mutate(drop = replace(drop, drop=="triangular_prism", "trig prism"),
         target = replace(target, target=="triangular_prism", "trig prism"))

@@

softmax <- function(weights, beta=1) {
  num <- exp(beta * weights)
  return(num / sum(num))
}

@@

evalFit <- function(feat_df, hum_df, beta) {
  hum_df <- get_model_target_choice_props(feat_df, hum_df, beta)
  r = with(hum_df, cor(chose_target, Model_targ_prop))
  return(-r) # return negative cor since DEoptim minimizes
}

@@

feat_names = names(mdat)[1:(ncol(mdat)-3)]
# for each feature, want to fit beta to maximize correlation between human and model choices
#feat_df <- get_feature_df(mdat, feat_names[1])

adult_fits = list()
for(feat in feat_names) {
  feat_df <- get_feature_df(mdat, feat)
  adult_fits[[feat]] = DEoptim(evalFit, lower=.01, upper=2, DEoptim.control(reltol=.001, NP=20, itermax=50), 
                feat_df=feat_df, hum_df=adult_trial_agg)
}
```

@@

child_fits = list()
for(feat in feat_names) {
  feat_df <- get_feature_df(mdat, feat)
  adult_fits[[feat]] = DEoptim(evalFit, lower=.01, upper=2, DEoptim.control(reltol=.001, NP=20, itermax=50), 
                feat_df=feat_df, hum_df=child_trial_agg)
}
```

>>>>

mdat <- mdat %>% rename(drop = drop_object,
                        target = target_object,
                        relation = condition) %>%
  mutate(drop = replace(drop, drop=="triangular_prism", "trig prism"),
         target = replace(target, target=="triangular_prism", "trig prism")) 
  #select(-`obj_final_position_std_objects=drop`, -`obj_final_position_std_objects=target`)
# these cols are all 0...

@@

softmax <- function(weights, beta=1, smooth=1e-7) {
  weights = weights + smooth
  probs = (weights / sum(weights)) # normalize first
  num <- exp(beta * probs)
  return(num / sum(num))
}

@@

evalFit <- function(feat_df, hum_df, beta) {
  hum_df <- get_model_target_choice_props(feat_df, hum_df, beta)
  #r = with(hum_df, cor(chose_target, Model_targ_prop))
  mse = with(hum_df, sum((chose_target - Model_targ_prop)^2)) / nrow(hum_df)
  return(mse) # return negative cor since DEoptim minimizes
}

@@

feat_names = names(mdat)[1:(ncol(mdat)-4)]
# for each feature, want to fit beta to maximize correlation between human and model choices
#feat_df <- get_feature_df(mdat, feat_names[1])

adult_fits = list()
for(feat in feat_names) {
  feat_df <- get_feature_df(mdat, feat)
  #paste(feat, evalFit(feat_df, adult_trial_agg, 1))
  adult_fits[[feat]] = DEoptim(evalFit, lower=.5, upper=5, DEoptim.control(reltol=.001, NP=20, itermax=50), 
                feat_df=feat_df, hum_df=adult_trial_agg)
}
```

@@

child_fits = list()
for(feat in feat_names) {
  feat_df <- get_feature_df(mdat, feat)
  child_fits[[feat]] = DEoptim(evalFit, lower=.5, upper=5, DEoptim.control(reltol=.001, NP=20, itermax=50), 
                feat_df=feat_df, hum_df=child_trial_agg)
}
```

```{r}
adult_mfit <- tibble()
child_mfit <- tibble()

for(f in names(adult_fits)) {
  adult_mfit = bind_rows(adult_mfit, c(feature=f, 
                                       beta=as.numeric(unlist(adult_fits[[f]]$optim$bestmem)), 
                                       mse=adult_fits[[f]]$optim$bestval))
  child_mfit = bind_rows(child_mfit, c(feature=f, 
                                       beta=as.numeric(unlist(child_fits[[f]]$optim$bestmem)), 
                                       mse=child_fits[[f]]$optim$bestval))
}

adult_mfit <- adult_mfit %>% 
  mutate(beta = as.numeric(beta),
         mse = as.numeric(mse)) %>%
  arrange(mse)

```

```{r children}
child_mfit <- child_mfit %>% 
  mutate(beta = as.numeric(beta),
         mse = as.numeric(mse)) %>%
  arrange(mse)

fit_tab <- child_mfit %>% rename(child_mse = mse, child_beta = beta) %>%
  left_join(adult_mfit %>% rename(adult_mse = mse, adult_beta = beta)) %>% 
  arrange(adult_mse)

knitr::kable(fit_tab, digits = 2)
```

```{r}
fits <- child_mfit %>% mutate(AgeGroup = "children") %>% 
  bind_rows(adult_mfit %>% mutate(AgeGroup = "adults"))

save(mdat, fits, file=here("models/model_data.RData"))
```

```{r, eval=F}
fits %>% ggplot(aes(x=beta, y=mse, color=AgeGroup)) +
  #geom_text(aes(label=feature)) + 
  geom_point() +
  theme_minimal()
```
So generally, children's choices are best explained by small values of beta (.5, the minimum tested).

What do the predicted choice proportions from the best feature for each age group look like?

```{r}
require(ggpubr)

plot_model_preds <- function(feat_name, fits, hum_data, age="adults") {
  beta = subset(fits, AgeGroup==age & feature==feat_name)$beta
  md <- get_model_target_choice_props(get_feature_df(mdat, feat_name), hum_data, beta)
  md %>% ggplot(aes(x=chose_target, y=Model_targ_prop, color=relation)) +
    geom_abline(slope=1, intercept=0, linetype="dashed") +
    geom_point(alpha=.5) + 
    xlab("Human Choice Probability") + 
    ylab("Model Choice Probability") +
    ggtitle(paste0(feat_name,": ",age)) + theme_minimal() + 
    xlim(0,1) + ylim(0,1)
}

ad_supp <- plot_model_preds("support_probability", fits, adult_trial_agg, age="adults")
ch_supp <- plot_model_preds("support_probability", fits, child_trial_agg, age="children")

ggarrange(ad_supp, ch_supp, nrow=1, common.legend = T)
```

```{r}
ad <- plot_model_preds("sharpness_of_support_posjitter_response", fits, adult_trial_agg, age="adults")
ch <- plot_model_preds("sharpness_of_support_posjitter_response", fits, child_trial_agg, age="children")

ggarrange(ad, ch, nrow=1, common.legend = T)
```

```{r}
ad <- plot_model_preds("support_std", fits, adult_trial_agg, age="adults")
ch <- plot_model_preds("support_std", fits, child_trial_agg, age="children")

ggarrange(ad, ch, nrow=1, common.legend = T)
```

```{r}
ad <- plot_model_preds("len_inverse_sharpe_ratio", fits, adult_trial_agg, age="adults")
ch <- plot_model_preds("len_inverse_sharpe_ratio", fits, child_trial_agg, age="children")

ggarrange(ad, ch, nrow=1, common.legend = T)
```
